Pulling oral and IV data from CvTdb

### 2025-05-06 data-raw/pulling_oral_iv_data_cvtdb.Rmd
Updated CvTdb SQL query and helper function.

  query_result <- tryCatch(
    return(dbGetQuery(con, query %>%
                        gsub("FROM ", paste0("FROM ", schema, "."), .) %>%
                        gsub("JOIN ", paste0("JOIN ", schema, "."), .))),
  error = function(cond){ 
    cat(paste0("\nError message: ", cond))
    return(NULL)
  },
  finally = { 
    dbDisconnect(con)
    }
    )
  return(query_result)
}

@@

```{r SQL_queries, warning=FALSE}
# Need to make this in a specified ordered
# First all IDs (so I can check them quickly)
# Note that this uses an ALIAS for each table which is later specified in
# the FROM and LEFT JOIN statments
query <- paste0(
  "SELECT distinct ",
  "a.id, a.time_original,  a.time_hr, a.conc_original, a.conc_sd_original, a.conc, a.conc_sd, ",
  #
  "a.fk_series_id, b.analyte_name_original, b.analyte_dtxsid, b.analyte_casrn, ",
  "b.fk_analyzed_chemical_id, b.fk_test_chemical_id, ",
  "b.test_substance_dtxsid, b.analyte_name_secondary_original, ",
  "b.conc_medium_normalized, b.conc_medium_original, b.time_units_original, b.conc_units_original, ",
  "b.conc_units_normalized, b.loq, b.loq_units, b.n_subjects_in_series, b.radiolabeled, ",
  #
  "b.fk_study_id, c.administration_route_normalized, c.fk_dosed_chemical_id, ",
  "c.dose_volume, c.dose_volume_units, c.dose_vehicle, ",
  "c.dose_duration, c.dose_duration_units, c.dose_frequency, c.fasting_period, ",
  "c.dose_level_normalized, c.dose_level_original, c.dose_level_units_original, ",
  #
  "b.fk_subject_id, d.weight_kg, d.species, d.sex, d.age, d.age_units, d.age_category, ",
  #
  "c.fk_extraction_document_id, e.pmid, e.year, e.other_study_identifier, ",
  "e.url, e.doi, e.extracted, e.curation_set_tag ",

@@

  # Filtering steps
  "WHERE c.administration_route_normalized IN ('iv', 'oral') AND ",
  "(b.conc_medium_normalized IN ('blood', 'plasma') OR ",
  "b.conc_medium_original IN ('blood', 'plasma'))"
)

>>>>

  query_result <- tryCatch(
    return(dbGetQuery(con, query %>%
                        gsub("FROM ", paste0("FROM ", schema, "."), .) %>%
                        gsub("JOIN (?=\\w)", paste0("JOIN ", schema, "."), ., perl = TRUE))),
  error = function(cond){ 
    cat(paste0("\nError message: ", cond))
    return(NULL)
  },
  finally = { 
    dbDisconnect(con)
    }
    )
  return(query_result)
}

@@

```{r SQL_queries, warning=FALSE}
# Need to make this in a specified ordered
# First all IDs (so I can check them quickly)
# Note that this uses an ALIAS for each table which is later specified in
# the FROM and LEFT JOIN statments
query <- paste0(
  "SELECT distinct ",
  # Conc_Time_Values table fields
  "a.id, a.fk_series_id, a.time_original, a.time_hr, ",
  "a.conc_original, a.conc_sd_original, a.conc, a.conc_sd, ",

  # Series table fields
  "b.fk_analyzed_chemical_id, ",
  ## Chemical dictionary fields (analyzed chemical information)
  "l.analyzed_chem_dtxsid, l.analyzed_chem_name_original, ",
  "l.analyzed_chem_casrn, l.analyzed_chem_name, ",

  "b.time_units_original, b.conc_units_original, ",
  "b.conc_units_normalized, b.loq, b.loq_units, ",
  "b.n_subjects_in_series, b.radiolabeled, b.fk_study_id, ",

  # Studies table fields
  "c.fk_dosed_chemical_id, ",
  ## Chemical dictionary fields (dosed chemical information)
  "k.dosed_chem_dtxsid, k.dosed_chem_name_original, ",
  "k.dosed_chem_casrn, k.dosed_chem_name, ",

  "c.dose_volume, c.dose_volume_units, c.dose_vehicle, ",
  "c.dose_duration, c.dose_duration_units, ",
  ## dose_frequency dictionary fields
  "j.dose_frequency_original, j.dose_frequency_normalized, ",
  "c.fasting_period, ",
  "c.dose_level_normalized, c.dose_level_original, c.dose_level_units_original, ",
  ## conc_medium dictionary fields
  "i.conc_medium_original, i.conc_medium_normalized, ", 
  ## administration_route dictionary fields
  "h.administration_route_original, h.administration_route_normalized, ",
  ## administration_method dictionary fields
  "g.administration_method_original, g.administration_method_normalized, ",
  ## administration_form dictionary fields
  "f.administration_form_original, f.administration_form_normalized, ",

  # Subjects table fields
  "b.fk_subject_id, d.weight_kg, d.species, d.sex, d.age, d.age_units, d.age_category, ",
  # Documents table fields
  "c.fk_extraction_document_id, e.pmid, e.year, e.other_study_identifier, ",
  "e.url, e.doi, e.extracted, e.curation_set_tag ",

@@

  # Join to dictionary tables
  "LEFT JOIN administration_form_dict f ON c.fk_administration_form_id = f.id ",
  "LEFT JOIN administration_method_dict g ON c.fk_administration_method_id = g.id ",
  "LEFT JOIN administration_route_dict h ON c.fk_administration_route_id = h.id ",
  "LEFT JOIN conc_medium_dict i ON b.fk_conc_medium_id = i.id ",
  "LEFT JOIN dose_frequency_dict j ON c.fk_dose_frequency_id = j.id ",

  # Rename chemical fields for dosed vs. analyzed chemical record foreign keys
  "LEFT JOIN (SELECT id, dsstox_substance_id as dosed_chem_dtxsid, ",
  "chemical_name_original as dosed_chem_name_original, ", 
  "dsstox_casrn as dosed_chem_casrn, preferred_name as dosed_chem_name ",
  "FROM chemicals) as k ON c.fk_dosed_chemical_id = k.id ",
  "LEFT JOIN (SELECT id, dsstox_substance_id as analyzed_chem_dtxsid, ",
  "chemical_name_original as analyzed_chem_name_original, ",
  "dsstox_casrn as analyzed_chem_casrn, preferred_name as analyzed_chem_name ",
  "FROM chemicals) as l ON b.fk_analyzed_chemical_id = l.id ",

  # Filtering steps
  "WHERE h.administration_route_normalized IN ('iv', 'oral') AND ",
  "(i.conc_medium_normalized IN ('blood', 'plasma', 'excreta') OR ",
  "i.conc_medium_original IN ('blood', 'plasma', 'excreta'))"
)
